---
title: "EISA proteomics - Create  database"
output: html_notebook
---

```{r libraries, message=FALSE, warning=FALSE, results=FALSE, echo=FALSE}
library(here)
library(tidyverse)
library(readxl)
library(openxlsx)
library(OrgMassSpecR)

source("./PepIsotopeDistribution.R")
```

```{r geneneral_stuff}
work_dir <- here()
data_dir <- here(work_dir, "Data")

db_file <- "BSA_peptides.xlsx"
```

# Read excel file

* Sheet1 contains a list with the precursors.
* Sheet3 contains for each peptide the b and y fragments

```{r results=FALSE}
precursors_info <- read_xlsx(path = here(data_dir, db_file),
                             sheet = "Sheet1") %>% 
  # remove some empty columns
  select(-c(`...9`, `...10`, `...11`))

# clean up
peptide_info <- precursors_info  %>% 
  # rename some columns
  rename(peptide = `peptides (observed)`,
         mz_prec = `M+H`,
         mz_2 = `2+`,
         mz_3 = `3+`) %>% 
  mutate(
    # transform the retention time to real numbers instead of characters
    RT = as.numeric(str_extract(string = RT,
                                pattern = "[0-9]{1,2}\\.?[0-9]{1,2}")),
    pep_length = as.numeric(str_extract(string = peptide,
                                        pattern = "[0-9]*$")) -
      as.numeric(str_extract(string = peptide,
                             pattern = "^[0-9]*")) + 1,
    pep_id = 1:n()
  ) %>% 
  relocate(pep_id, .before = peptide)
```

There are 38 peptides in the Excel sheet.

Define the start position of each peptide.

```{r}
range_fragments <- tibble(pep_id = 1:38,
                          start_row = c(1, 14, 25, 39, 55, 69,
                                        84, 103, 114, 130, 139, 158,
                                        175, 192, 205, 218, 237, 253,
                                        270, 285, 299, 321, 338, 354,
                                        372, 388, 406, 426, 446, 466,
                                        480, 501, 523, 539, 557, 578,
                                        601, 618) # this is including the header of each table
)
```

Read all the tables from `Sheet3`.

```{r results=FALSE}
all_info <- peptide_info %>% 
  left_join(y = range_fragments,
            by = "pep_id") %>% 
  mutate(start_range = paste0("A", start_row, ":S", start_row + pep_length),
         # read all fragment info
         fragments = map(.x = start_range,
                         .f = ~ read_xlsx(path = here(data_dir, db_file),
                                          sheet = "Sheet3",
                                          range = .x) %>% 
                           # add numbering of fragments
                           mutate(b_num = 1:n(),
                                  y_num = n():1) %>% 
                           # ony select the columns which I want to use
                           select(b_num, b, y_num, y) %>% 
                           # make longer
                           pivot_longer(cols = c(b, y),
                                        names_to = "fragment",
                                        values_to = "mz_fragment") %>% 
                           # create fragment names
                           mutate(fragment = if_else(fragment == "b",
                                              paste0(fragment, b_num),
                                              paste0(fragment, y_num))) %>% 
                           # keep only the columns I want
                           select(fragment, mz_fragment))) %>% 
  # unfold
  unnest(fragments) %>% 
  pivot_longer(cols = c(mz_prec, mz_fragment),
               names_to = "type",
               values_to = "mz") %>% 
  select(pep_id, peptide, sequence, type, Mass, mz, mz_2, mz_3, fragment, RT) %>% 
  filter(!is.na(mz)) %>% 
  mutate(Mass = if_else(type == "mz_prec",
                        Mass,
                        NA_real_),
         mz_3 = if_else(type == "mz_prec",
                        mz_3,
                        NA_real_),
         mz_2 = if_else(type == "mz_prec",
                        mz_2,
                        mz / 2 + 0.5037),
         fragment = if_else(type == "mz_prec",
                            "",
                            fragment)) %>% 
  distinct(sequence, mz, .keep_all = TRUE) %>% 
  # create charge column
  rename(mz_1 = mz) %>% 
  pivot_longer(cols = starts_with("mz"),
               names_to = "charge",
               values_to = "mz",
               names_prefix = "mz_",
               names_transform = list(charge = as.integer)) %>% 
  mutate(fragment = if_else(type == "mz_prec",
                            "precursor",
                            fragment)) %>% 
  filter(!is.na(mz)) %>% 
  select(pep_id, peptide, sequence, Mass, RT, fragment, mz, charge)
```

# Calculate theoretical isotopic distributions. This takes a long time to compute.

```{r}
all_info <- all_info %>% 
  mutate(series = if_else(fragment != "precursor",
                          str_extract(string = fragment,
                                      pattern = "[by]"),
                          ""),
         ms2type = if_else(fragment == "precursor",
                          "",
                          paste0("[", fragment, "]", charge, "+")),
         # get the sequence for the fragments
         ms2seq = map2_chr(.x = sequence,
                           .y = ms2type,
                           .f = ~ ifelse(.y == "",
                                         "",
                                         FragmentPeptide(sequence = .x) %>% 
                                           filter(ms2type == .y) %>% 
                                           pull(ms2seq))),
         ms2seq = ifelse(is.na(ms2seq),
                         sequence,
                         ms2seq),
         # calculate the theoretical distriution
         theor_isotopes = pmap(.l = list(ms2seq, charge, series),
                               .f = ~ PepIsotopeDistribution(sequence = ..1,
                                                             charge = ..2,
                                                             series = ..3) %>%
                                 select(mz, percent)))
```

# Export 

Export to new Excel file.

```{r eval=FALSE}
write.xlsx(x = all_info,
           file = here(data_dir, "peptide_db.xlsx"),
           overwrite = TRUE)
```

This data base cannot be exported to Excel. I will export it to an .Rdata file.

```{r}
db_data <- all_info
save(db_data,
     file = here(data_dir, "peptide_db.Rdata"))
```



```{r session_info}
sessioninfo::session_info()
```

