---
title: "EISA proteomics"
output:
  html_notebook:
    code_folding: hide
---

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
library(here)
library(tidyverse)
library(xcms)
library(readxl)
library(ggrepel)
library(ggforce)
library(foreach)
library(doParallel)
library(OrgMassSpecR)
library(ggrepel)

source("./PepIsotopeDistribution.R")
```

```{r geneneral_stuff}
work_dir <- here()
data_dir <- here(work_dir, "Data")

data_file <- "BSA_eISA_redo_50_60eV_p1a4_01_8086.mzML"
db_file <- "peptide_db.Rdata"

# +/- seconds
rt_window <- 10
# +/- Da
mz_window <- 0.01
# set the correlatoin coefficient limit
cor_limit <- 0.8
```

# Read the MS file

The mzML file contains MS levels. MS level 1 is MS1 and MS level 2 is the isCID experiment.

```{r}
raw_data <- readMSData(files = here(data_dir, data_file),
                       msLevel. = 2,
                       mode = "onDisk")
```

Show the BPC of the isCID experiment.

```{r}
bpis <- chromatogram(object = raw_data, 
                     aggregationFun = "max",
                     msLevel = 2L)

plot(bpis)
```

# Read the database

```{r}
load(file = here(data_dir, db_file))
```

Calculate isotopes. ***Not sure it this is still needed***

```{r}
db_data <- db_data %>% 
  mutate(M1 = mz + 1.003 / charge,
         M2 = mz + (2 * 1.003) / charge,
         M3 = mz + (3 * 1.003) / charge)
```

Show the first few lines of the database.

```{r}
db_data %>% 
  head(n = 10)
```

Calculate isotopes.

```{r}
db_data <- db_data %>% 
  mutate(M1 = mz + 1.003 / charge,
         M2 = mz + (2 * 1.003) / charge,
         M3 = mz + (3 * 1.003) / charge,
         M4 = mz + (4 * 1.003) / charge)
```

# Analysis

For each peptide create extracted ion chromatograms for all precursor ions en fragment ions.

# One peptide

Start with one peptide (SLHTLFGDELCK).

```{r}
selected_pep_id <- 1

peptide <- db_data %>% 
  filter(pep_id == selected_pep_id) %>% 
  mutate(mz_char = sprintf("%.4f", mz))

mzs <- matrix(c(peptide$mz - mz_window, peptide$mz + mz_window), ncol = 2)
rts <- unique(peptide$RT) * 60 + c(-rt_window, rt_window)

eics <- chromatogram(object = raw_data,
                     mz = mzs,
                     rt = rts,
                     msLevel = 2L,
                     aggregationFun = "max")
```

Remove all EIC's which have more than 80% NA's.

Show all EIC's. 

```{r}
# put everything in a single data frame
eic_df_wide <- cbind(tibble(rt = rtime(eics[[1]])), mapply(intensity, eics))
# set the column names
colnames(eic_df_wide) <- c("rt", sprintf("%.4f", peptide$mz))

# calculate the ratio of NA's
num_na <- sapply(eic_df_wide[, -1], function(x) mean(is.na(x)))
num_na_names <- names(which(num_na > 0.8))

# make long format
eic_df <- eic_df_wide %>% 
  # remove columns (EIC's) which contain too many NA's
  select(-any_of(num_na_names)) %>% 
  pivot_longer(cols = -rt,
               names_to = "eic",
               values_to = "intens") %>% 
  group_by(eic) %>% 
  mutate(show_eic = ifelse(intens == max(intens, na.rm = TRUE),
                           eic,
                           NA_character_))
```

```{r fig.width=10, warning=FALSE}
eic_df %>% 
  ggplot(aes(x = rt,
             y = intens)) +
  geom_line(aes(color = eic,
                group = eic)) +
  # geom_text_repel(aes(label = show_eic)) +
  guides(color = "none") +
  labs(x = "Retention time [s]",
       y = "Intensity",
       title = paste0("Peptide: ", db_data$sequence[db_data$pep_id == 5][selected_pep_id])) +
  facet_zoom(y = intens < 1e4,
             zoom.size = 1) +
  theme_minimal()
```

```{r fig.height=20, fig.width=10, warning=FALSE}
eic_labels <- paste0(peptide$fragment, "(", peptide$charge, "+) - ", peptide$mz_char)
names(eic_labels) <- peptide$mz_char

eic_df %>% 
  left_join(y = peptide,
            by = c("eic" = "mz_char")) %>% 
  ggplot(aes(x = rt,
             y = intens)) +
  geom_line(aes(color = eic,
                group = eic)) +
  geom_vline(xintercept = peptide$RT[selected_pep_id] * 60,
             linetype = 2) +
  guides(color = "none") +
  scale_y_continuous(labels = scales::scientific) +
  labs(x = "Retention time [s]",
       y = "Intensity",
       title = paste0("Peptide: ", db_data$sequence[db_data$pep_id == 5][selected_pep_id])) +
  facet_wrap(vars(eic),
             scales = "free_y",
             ncol = 3,
             labeller = labeller(eic = eic_labels)) +
  theme_minimal()
```

Determine which of the precursor signals (1+, 2+ or 3+) is the highest.

```{r}
# get the highest value for each precursor singal
high_precursor <- sapply(eic_df_wide[, 2:4], max, na.rm = TRUE)
highest_precursor <- names(which(high_precursor == max(high_precursor)))

highest_precursor
```

Get the correlation of all EIC's to this precursor EIC.

```{r}
# remove the columns which contain too many NA's
eic_cor_df <- eic_df_wide[, -1] %>% 
  select(-any_of(num_na_names))
# replace NA's by zero
eic_cor_df[is.na(eic_cor_df)] <- 0

# calculate the correlation between the EIC's
eic_cor_result <- cor(eic_cor_df)[, highest_precursor]
```

Show all correlation coeffiencts. 710.3505 is the precursor ion with the highest signal.

```{r}
# transform to dataframe, first 3 rows are the precursor signals
cor_result_df <- as_tibble(eic_cor_result,
                           rownames = "fragment_mz") 

cor_result_df %>% 
  arrange(desc(value))
```

The fragment percentage found.

```{r}
peptide_res <- peptide %>% 
  left_join(y = cor_result_df,
            by = c("mz_char" = "fragment_mz")) %>% 
  mutate(found = ifelse(value >= cor_limit,
                        TRUE,
                        FALSE),
         found = ifelse(is.na(found),
                        FALSE,
                        found))

peptide_res %>% 
  filter(fragment != "precursor") %>% 
  group_by(series, charge) %>% 
  summarise(frag_found = mean(found),
            .groups = "drop")
```

## Isotopic distribution

Get all raw MS spectra.

```{r eval=FALSE}
raw_df <- as(raw_data, "data.frame")
```

Extract spectrum at the peak top.

```{r eval=FALSE}
pep_res_filtered <- peptide_res %>% 
  mutate(spec = pmap(.l = list(RT, mz, charge),
                     .f = ~ raw_df %>% 
                       filter(rt > ..1 * 60 - 0.25 &
                                rt < ..1 * 60 + 0.25,
                              mz > ..2 - 0.2 &
                                mz < ..2 + 4 / ..3) %>% 
                       mutate(i_perc = i / max(i) * 100) %>% 
                       select(mz, i_perc)))
```

Calculate the spectrum similarity.

```{r eval=FALSE}
pep_res_filtered <- pep_res_filtered %>% 
  mutate(similarity = map2_dbl(.x = spec,
                               .y = theor_isotopes,
                               .f = ~ SpectrumSimilarity(spec.top = .x,
                                                         spec.bottom = .y,
                                                         t = 0.01,
                                                         print.graphic = FALSE)))
```

Alternative method. Get the maximimum of each of the EIC's of the isotopes and create a spectrum. For now only M+0, M+1, M+2 and M+3 are used.

```{r results='hide'}
# get the EIC's of mz, M1, M2 and M3
get_measured_isotopes <- function(M0, M1, M2, M3, RT) {
  mzs <- matrix(c(M0 - mz_window, M0 + mz_window,
                  M1 - mz_window, M1 + mz_window + mz_window,
                  M2 - mz_window, M2 + mz_window + mz_window,
                  M3 - mz_window, M3 + mz_window + mz_window),
                ncol = 2, byrow = TRUE)
  rts <- unique(peptide$RT) * 60 + c(-rt_window, rt_window)
  
  eics <- chromatogram(object = raw_data,
                       mz = mzs,
                       rt = rts,
                       msLevel = 2L,
                       aggregationFun = "max")
  
  max_intens <- apply(mapply(intensity, eics), 2, function(x) max(x, na.rm = TRUE))
  
  res <- data.frame(mz = c(M0, M1, M2, M3),
                    i = max_intens) %>% 
    mutate(i = ifelse(is.infinite(i),
                      0, 
                      i),
           i_perc = i / max(i) * 100,
           i_perc = ifelse(is.nan(i_perc),
                           0,
                           i_perc)) %>% 
    select(-i)
  
  return(res)
}

peptide_res <- peptide_res %>% 
  mutate(meas_isotopes = pmap(.l = list(mz, M1, M2, M3, RT),
                              .f = ~ get_measured_isotopes(M0 = ..1,
                                                           M1 = ..2,
                                                           M2 = ..3,
                                                           M3 = ..4,
                                                           RT = ..5)))

```

Calculate similarity.

```{r}
peptide_res <- peptide_res %>% 
  mutate(similarity = map2_dbl(.x = meas_isotopes,
                               .y = theor_isotopes,
                               .f = ~ SpectrumSimilarity(spec.top = .x,
                                                         spec.bottom = .y,
                                                         t = 0.025,
                                                         b = 5,
                                                         print.graphic = FALSE)))
```

The fragment percentage, with isotope distribution filtering.

```{r}
peptide_res <- peptide_res %>% 
  mutate(found = ifelse(value >= cor_limit &
                          similarity >= 0.75,
                        TRUE,
                        FALSE),
         found = ifelse(is.na(found),
                        FALSE,
                        found))

peptide_res %>% 
  filter(fragment != "precursor") %>% 
  group_by(series, charge) %>% 
  summarise(frag_found = mean(found),
            .groups = "drop")
```

```{r}
# if 3rd or 4th isotope is 100% remove them
cleanup_isotopes <- function(x) {
  if(x$i_perc[3] == 100 | x$i_perc[4] == 100) {
    res <- x[-c(3, 4), ]
  } else {
    res <- x
  }
  return(res)
}

peptide_res <- peptide_res %>%
  mutate(meas_isotopes = map(.x = meas_isotopes,
                             .f = ~ cleanup_isotopes(.x)))

peptide_res <- peptide_res %>% 
  mutate(similarity = map2_dbl(.x = meas_isotopes,
                               .y = theor_isotopes,
                               .f = ~ SpectrumSimilarity(spec.top = .x,
                                                         spec.bottom = .y,
                                                         t = 0.025,
                                                         b = 5,
                                                         print.graphic = FALSE)))

peptide_res <- peptide_res %>% 
  mutate(found = ifelse(value >= cor_limit &
                          similarity >= 0.75,
                        TRUE,
                        FALSE),
         found = ifelse(is.na(found),
                        FALSE,
                        found))

peptide_res %>% 
  filter(fragment != "precursor") %>% 
  group_by(series, charge) %>% 
  summarise(frag_found = mean(found),
            .groups = "drop")
```


# All peptides

Do all peptides at once.

```{r eval=FALSE}
# start the cluster
cl <- makeCluster(detectCores()[1] - 2) #not to overload your computer
registerDoParallel(cl)

rt_windows <- c(5, 10, 15, 20, 25, 30, 35, 40)
rt_all_results <- NULL

for (rt_window in rt_windows) {
  
  all_results <- foreach (select_pep_id = 1:38, 
                          .combine = rbind, 
                          .packages = c("tidyverse", "xcms", "OrgMassSpecR")) %dopar% {
                            # get the info of one peptide
                            peptide <- db_data %>% 
                              filter(pep_id == select_pep_id) %>% 
                              mutate(mz_char = sprintf("%.4f", mz))
                            
                            # create mz and rt ranges
                            mzs <- matrix(c(peptide$mz - mz_window, peptide$mz + mz_window), ncol = 2)
                            rts <- unique(peptide$RT) * 60 + c(-rt_window, rt_window)
                            
                            # get all EIC's
                            eics <- chromatogram(object = raw_data,
                                                 mz = mzs,
                                                 rt = rts,
                                                 msLevel = 2L,
                                                 aggregationFun = "max")
                            ### correlation part
                            # put everything in a single data frame
                            eic_df_wide <- cbind(tibble(rt = rtime(eics[[1]])), mapply(intensity, eics))
                            # set the column names
                            colnames(eic_df_wide) <- c("rt", sprintf("%.4f", peptide$mz))
                            
                            # calculate the ratio of NA's
                            num_na <- sapply(eic_df_wide[, -1], function(x) mean(is.na(x)))
                            num_na_names <- names(which(num_na > 0.8))
                            
                            # get the highest value for each precursor signal
                            high_precursor <- sapply(eic_df_wide[, 2:4], max, na.rm = TRUE)
                            highest_precursor <- names(which(high_precursor == max(high_precursor, na.rm = TRUE)))
                            
                            # remove the columns which contain too many NA's
                            eic_cor_df <- eic_df_wide[, -1] %>% 
                              select(-any_of(num_na_names))
                            # replace NA's by zero
                            eic_cor_df[is.na(eic_cor_df)] <- 0
                            
                            # calculate the correlation between the EIC's
                            eic_cor_result <- cor(eic_cor_df)[, highest_precursor]
                            
                            # transform to dataframe, first 3 rows are the precursor signals
                            cor_result_df <- as_tibble(eic_cor_result,
                                                       rownames = "fragment_mz")
                            
                            peptide_res <- peptide %>% 
                              left_join(y = cor_result_df,
                                        by = c("mz_char" = "fragment_mz")) %>% 
                              mutate(found = ifelse(value >= cor_limit,
                                                    TRUE,
                                                    FALSE),
                                     found = ifelse(is.na(found),
                                                    FALSE,
                                                    found),
                                     rt_window = 2 * rt_window,
                                     # get the measure isotope ratio
                                     meas_isotopes = pmap(.l = list(mz, M1, M2, M3, RT),
                                                          .f = ~ get_measured_isotopes(M0 = ..1,
                                                                                       M1 = ..2,
                                                                                       M2 = ..3,
                                                                                       M3 = ..4,
                                                                                       RT = ..5)),
                                     # if 3rd or 4th isotope is 100%, remove them, this due to overlapping ion
                                     meas_isotopes = map(.x = meas_isotopes,
                                                         .f = ~ cleanup_isotopes(.x)))
                            
                            #  calculate the similarity
                            peptide_res <- peptide_res %>% 
                              mutate(similarity = map2_dbl(.x = meas_isotopes,
                                                           .y = theor_isotopes,
                                                           .f = ~ SpectrumSimilarity(spec.top = .x,
                                                                                     spec.bottom = .y,
                                                                                     t = 0.025,
                                                                                     b = 5,
                                                                                     print.graphic = FALSE)))
                            
                            peptide_res
                          }
  
  rt_all_results <- rbind(rt_all_results, all_results)
}
# stop the cluster
stopCluster(cl)

save(rt_all_results,
     file = here(data_dir, "all_results.Rdata"))
```


```{r eval=FALSE}
load(here(data_dir, "all_results.Rdata"))
```

## No filtering isotope distribution

The median of the fragment percentage for retention time window 30s (+/- 15s). No filtering based on isotope distribution.

```{r}
# calculate the fragment found percentage
frag_found <- rt_all_results %>% 
  filter(fragment != "precursor") %>% 
  group_by(rt_window, pep_id, series, charge) %>% 
  summarise(frag_found = mean(found, na.rm = TRUE),
            pep_id = pep_id[1],
            rt_window = rt_window[1],
            .groups = "drop") 

frag_found %>%
  filter(rt_window == 30) %>% 
  group_by(series, charge) %>% 
  summarise(median = median(frag_found),
            .groups = "drop")
```

Show the fragment percentages for all retention time windows.

```{r fig.width=8, fig.height=6}
frag_found %>% 
  mutate(frag_found = frag_found * 100,
         series = paste0(series, charge, "+"),
         rt_window = factor(rt_window)) %>% 
  ggplot(aes(x = series,
             y = frag_found,
             fill = rt_window)) +
  geom_boxplot(aes()) +
  # geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  # geom_point(size = 1,
  #            position=position_jitterdodge(jitter.width = 0.1)) +
  labs(x = "Fragment ion series",
       y = "Fragment percentage [%]",
       fill = "RT window") +
  # coord_flip() +
  theme_minimal()
```

```{r fig.width=8, fig.height=6}
frag_found %>% 
  mutate(frag_found = frag_found * 100,
         series = paste0(series, charge, "+"),
         rt_window = factor(rt_window)) %>% 
  ggplot(aes(x = series,
             y = frag_found,
             fill = rt_window)) +
  geom_boxplot(aes()) +
  # geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  # geom_point(size = 1,
  #            position=position_jitterdodge(jitter.width = 0.1)) +
  labs(x = "Fragment ion series",
       y = "Fragment percentage [%]",
       fill = "RT window",
       title = "No Filtering Isotope Distribution") +
  # coord_flip() +
  theme_minimal()
```

## Filtering isotope distribution

If the isotope distribution similarity is not high enough (< 0.75) the fragment is set to not being found. The median of the fragment percentage for retention time window 30s (+/- 15s). With filtering based on isotope distribution similarity.

```{r}
# calculate the fragment found percentage
frag_found_iso <- rt_all_results %>% 
  filter(fragment != "precursor") %>%
  mutate(similarity = ifelse(is.nan(similarity),
                             0,
                             similarity),
         found = ifelse(value >= cor_limit &
                          similarity >= 0.75,
                        TRUE,
                        FALSE)) %>% 
  group_by(rt_window, pep_id, series, charge) %>% 
  # calculate fragment percentage found
  summarise(frag_found = mean(found, na.rm = TRUE),
            pep_id = pep_id[1],
            rt_window = rt_window[1],
            .groups = "drop") 

frag_found_iso %>%
  filter(rt_window == 30) %>% 
  group_by(series, charge) %>% 
  summarise(median = median(frag_found),
            .groups = "drop")
```

Show the fragment percentages for all retention time windows.

```{r fig.width=8, fig.height=6}
frag_found_iso %>% 
  mutate(frag_found = frag_found * 100,
         series = paste0(series, charge, "+"),
         rt_window = factor(rt_window)) %>% 
  ggplot(aes(x = series,
             y = frag_found,
             fill = rt_window)) +
  geom_boxplot(aes()) +
  # geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  # geom_point(size = 1,
  #            position=position_jitterdodge(jitter.width = 0.1)) +
  labs(x = "Fragment ion series",
       y = "Fragment percentage [%]",
       fill = "RT window") +
  # coord_flip() +
  theme_minimal()
```

```{r fig.width=8, fig.height=6}
frag_found_iso %>% 
  mutate(frag_found = frag_found * 100,
         series = paste0(series, charge, "+"),
         rt_window = factor(rt_window)) %>% 
  ggplot(aes(x = series,
             y = frag_found,
             fill = rt_window)) +
  geom_boxplot(aes()) +
  # geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  # geom_point(size = 1,
  #            position=position_jitterdodge(jitter.width = 0.1)) +
  labs(x = "Fragment ion series",
       y = "Fragment percentage [%]",
       fill = "RT window",
       title = "Filtering Isotope Distribution") +
  # coord_flip() +
  theme_minimal()
```

# Session info

```{r session_info}
sessioninfo::session_info()
```

